---
title: "Geog533 Lab 8"
author: "Josh Gonzalez"
output:
  html_document:
    toc: yes
  html_notebook:
    toc: yes
    toc_float: yes
---

Complete the following exercises in Chapter 8 (Introduction to Regression Analysis) of the textbook [R] pages 247-250. 

## Question 1

This is Exercise 1 in Chapter 8 of the Textbook [R].

A regression of weekly shopping trip frequency on annual income (data entered in thousands of dollars) is performed on data collected from 24 respondents. The results are summarized below:

* Intercept: 0.46
* Slope: 0.19

|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression|                 |                    |                |                 |
| Residual |       1.7       |                    |                |                 |
| Total    |       2.3       |                    |                |                 |


### Question 1(a)
Fill in the blanks in the ANOVA table.

```{r}
n <- 24
ssqr <- 2.3-1.7
dof <- n-2 
doftot <- n-1
msr0 <- 1.7/dof
msr1 <- round(msr0, 3)
f <- round(ssqr/msr0, 2)
f
```


|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression|      r 'ssqr'   |                    |     r 'ssqr'   |        r 'f'    |
| Residual |       1.7       |         r 'dof'    |                |                 |
| Total    |       2.3       |        r 'doftot'  |                |                 |


### Question 1(b)
What is the predicted number of weekly shopping trips for someone making $50,000/year?
```{r}
0.46+(0.19*50)
```


### Question 1(c)
In words, what is the meaning of the coefficient 0.19?
A person will go on .19 shopping trips per week for every dollar they make.


### Question 1(d)
Is the regression coefficient significantly different from zero? How do you know?
```{r}
c <- qf(.95,1,22)
if(f < c){
print("We cannot reject the null hypothesis")
} else {
print("We reject the null hypothesis")}
```
"This proves that the regression coefficient is signigicantly different than zero because the F value is greater than the critical F value"


### Question 1(e)
What is the value of the correlation coefficient?
```{r}
r <- sqrt(ssqr/2.3)
round(r,3)
```





## Question 2

This is Exercise 6 in Chapter 8 of the Textbook [R].

The following data are collected in an effort to determine whether snowfall is dependent upon elevation: 

| Snowfall (inches) | Elevation (feet) |
|-------------------|------------------|
|       36          |       400        |
|       78          |       800        |
|       11          |       200        |
|       45          |       675        |

Using R, show your work on exercises (a) through (g).

### Question 2(a)
Find the regression coefficients (the intercept and the slope coefficient).
```{r}
snow <- c(36,78,11,45)
elev <- c(400,800,200,675)
mn <- lm(snow~elev)
summary(mn)
int <- -7.63721
round(int, 3)
m <- 0.09665
round(m, 3)
```


### Question 2(b)
Estimate the standard error of the residuals about the regression line.
```{r}
print("109.8 for 2 degrees of Freedom.")
```


### Question 2(c)
Test the hypothesis that the regression coefficient associated with the independent variables is equal to zero. Also place a 95% confidence interval on the regression coefficient.
```{r}
f2 <- 16.7
c2 <- qf(.95, 1, 2)
if(f2 < c2){
print("We cannot reject the null hypothesis") 
} else {
print("We reject the null hypothesis")}
se <- 2.29
q95 <- qnorm(0.975)
c95 <- c(f2-q95*se, f2+q95*se)
c95
```


### Question 2(d)
Find the value of $r^2$.
```{r}
rsqrd <- 0.8899
apprsqrd <- 0.8349
```


### Question 2(e)
Make a table of the observed values, predicted values, and the residuals.
```{r}
dfse <- data.frame(snow,elev)
dfse$predicted <- round(mn$fitted.values,3)
dfse$residuals <- round(mn$residuals,3)
names(dfse) <- c("Snowfall (In)", "Elevation (Ft)", "Predicted Values", "Residuals")
library(knitr)
kable(dfse)
```


### Question 2(f)
Prepare an analysis of variance table portraying the regression results.
```{r}
anova(mn)
```


|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression|      2047.77    |        1           |   2047.77      |        16.17    |
| Residual |      253.23     |        2           |   126.61       |                 |
| Total    |      2301       |        3           |                |                 |

### Question 2(g)
Graph the data and the regression line.
```{r}
plot(elev, snow)
abline(mn, col = "blue", lwd = 2)
segments(elev,fitted(mn),elev,snow,col="purple",lty = "dashed")
```




## Question 3

This is Exercise 10 in Chapter 8 of the Textbook [R].

Use R and the [Milwaukee dataset](http://spatial.binghamton.edu/geog533/data/Milwaukee_Sales.csv) to:

### Problem
perform a regression using sales price as the dependent variable and lot size as the independent variable.

### Solution
```{r}
url = "http://spatial.binghamton.edu/geog533/data/Milwaukee_Sales.csv"
rbing <- read.csv(url)

site <- lm(rbing$SalePrice~rbing$LotSize)
summary(site)
```



## Question 4

This is Exercise 11 in Chapter 8 of the Textbook [R].

Use R and the Hypothetical [UK Housing Prices dataset](http://spatial.binghamton.edu/geog533/data/UK_Housing.csv) to:

### Question 4(a) 
perform a regression using house price as the dependent variable, and number of bedrooms as the independent variable;

```{r}
url = "http://spatial.binghamton.edu/geog533/data/UK_Housing.csv"
rbinguk <- read.csv(url)

siteuk <- lm(rbinguk$price~rbinguk$bedrooms)
summary(siteuk)
```


### Question 4(b) 
(b) repeat part (a), using the number of bathrooms as the independent variable, and comment on the results.
```{r}
bathind <- lm(rbinguk$price~rbinguk$bathrooms)

summary(bathind)
```


## Question 5
Download the [hometown.csv](http://spatial.binghamton.edu/geog533/data/hometown.csv), append your name and hometown coordinates to the csv, then use [leaflet](https://rstudio.github.io/leaflet/) to create a web map (10 pt.)

```{r}
dfbing <- read.csv("hometown.csv")
View(dfbing)
datame <-data.frame(lastname="Gonzalez", firstname="Joshua",type="student",location="40.7831,-73.9712",latitude = 40.7831,longitude = -73.9712,image="",fullname= "Joshua Gonzalez",popup="")
dfbing2 <- rbind(dfbing, datame)
View(dfbing2)

library(leaflet)
m <- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lng= -73.9712, lat= 40.7831, popup="Manhattan")
m  # Print the map
```

