---
title: "Geog533 Lab 10"
author: "Joshua Gonzalez"
output:
  html_document:
    toc: yes
  html_notebook:
    toc: yes
    toc_float: yes
---

## Question 1
Use R package UScensus2010county to complete the following tasks:  (20 pt.)
```{r}
if(!require(UScensus2010)) {install.packages("UScensus2010")}
library(UScensus2010)
if(!require(UScensus2010county)) {install.county("osx")}
library(UScensus2010county)
library(tmap)
```
### Question 1(a)
Plot a map of New York counties using the plot function.
```{r}
data(new_york.county10)
shpfile <- new_york.county10
df1 <- shpfile@data
df1
plot(shpfile)
```

### Question 1(b)	
Plot a map of New York counties using the qtm function.
```{r}
install.packages("tmap")
library(tmap)
qtm(shpfile)
```


### Question 1(c)	
How many counties in New York State?
```{r}
df
print("There are 62 counties in New York State")
```

### Question 1(d)	
What’s the 3-digit fips code of Broome County?
```{r}
print("36007")
```

### Question 1(e)	
Compute descriptive statistics of the population column (P0010001), including total, minimum, maximum, mean, median, and skewness. 
```{r}
library(moments)
nypop <- df$P0010001
sum(nypop)
min(nypop)
max(nypop)
mean(nypop)
median(nypop)
skewness(nypop)
```

### Question 1(f)	
Create a histogram and a boxplot of the population.
```{r}
a <- hist(nypop)
a
b <- boxplot(nypop)
b
```


## Question 2
Use R package UScensus2010tract to complete the following tasks:    (20 pt.)
```{r}
library(UScensus2010)
if(!require(UScensus2010tract)) install.tract("osx")
library(UScensus2010tract)
```

### Question 2(a)	
Plot a map of New York census tracts using the plot function.
```{r}
data(new_york.tract10)
shpfile <- new_york.tract10

df <- shpfile@data
df
plot(shpfile)
```

### Question 2(b)
Compute the total population based on census tracts.
```{r}
df
newpop <- df$P0010001

snew <- sum(newpop)
snew
```

### Question 2(c)
Select all census tracts in Broome County and plot the map. 
```{r}
broome <- county(fips="007",name="Broome",state="NY",level="tract")
plot(broome)
```

### Question 2(d)
What’s the total population of Broome County?
```{r}
bpop <- sum(broome$P0010001)
bpop
```

### Question 2(e)
Create a histogram and a boxplot of population based on census tracts of Broome County.
```{r}
hist(broome$P0010001)

boxplot(broome$P0010001)
```

### Question 2(f)
Select the first five columns of the shapefile of Broome County census tract; add a new population ratio column (= census tract population / county population); save the new shapefile to the hard drive. 
```{r}
popra <- data.frame(broome[,1:5])
popra$ratio <- broome$P0010001/200600
popra
```


## Question 3

Use R packages raster and ncdf4 to complete the following tasks:     (20 pt.)
```{r}
library(raster)
library(ncdf4)
```

### Question 3(a)		
Load the multi-band raster dataset “NDVI.nc” into R.
```{r}
ndvi <- raster("NDVI.nc")
```

### Question 3(b)		
Get the basic information about the dataset, including the number of rows, columns, cells, and bands; spatial resolution, extent, bounding box, and projection.
```{r}
nrow(ndvi)
ncol(ndvi)
ncell(ndvi)
nbands(ndvi)
res(ndvi)
extent(ndvi)
bbox(ndvi)
projection(ndvi)
```

### Question 3(c)	
Aggregate all bands to generate a mean NDVI raster and a maximum NDVI raster; save the two new raster datasets to the hard drive. 
```{r}
ndvi.rb <- brick("ndvi.nc")
meanndvi <- mean(ndvi.rb)
plot(meanndvi)
writeRaster(meanndvi,filename = "meanndvi.tif",overwrite=TRUE)
maxndvi <- max(ndvi.rb)
plot(maxndvi)
writeRaster(maxndvi,filename = "maxndvi.tif",overwrite=TRUE)
```

### Question 3(d)			
Plot the maps of monthly NDVI of the year 2001
```{r}
ndvi2k <- ndvi.rb[[13:24]]
plot(ndvi2k)
```

### Question 3(e)	
Create histograms of monthly NDVI of the year 2001.
```{r}
hist(ndvi2k)
```

### Question 3(f)			
Plot the NDVI map of July 2000; click any location with data on the map and retrieve the NDVI time series for all years; plot the NDVI time series of the selected location. 
```{r}
plot(ndvi.rb,7) 
location <- ndvi.rb[150, 50]
location <- click(ndvi.rb, n=1, xy=TRUE)
location <- click(ndvi.rb, n=1, xy=FALSE)
```

